Under-approximation of Reachability in Multivalued Asynchronous Networks

### 2013-04-02 parts/sa.tex
several fixes; first final version
minor fixes; third final version

\section{Static analysis}\label{sec:sa}

@@

We define $\concr(\w)$ as the set of scenarios concretising an objective sequence $\w$ in the context $\ctx$.
In \pref{def:uconcr}, we define $\uconcr(\w)$ as equal to $\concr(\w)$ if and only if $\concr(\w)$ contains scenarios starting from all states $s \subseteq \ctx$.
\pref{th:uconcr-ctx} is used to over-approximate the initial context $\ctx$.
\begin{definition}[$\uconcr: \OS \rightarrow \powerset(\Sce)$]
\label{def:uconcr}
  \[
  \uconcr(\w) = 
  \begin{cases}
    \concr(\w) & \text{if } \forall s \in \PHl, s \subseteq \ctx, \exists \delta \in \concr(\w), \delta \in \Sce(s) \\
    \emptyset & \text{else.}
  \end{cases}
  \]
\end{definition}
% 
\begin{theorem}
\label{th:uconcr-ctx}
  $\ctx \subseteq \ctx' \wedge \muconcr_{\ctx'}(\w) \neq \emptyset \Longrightarrow \muconcr_{\ctx}(\w) \neq \emptyset$.
\end{theorem}

@@

The graph of local causality $\cwB = (\V, \E)$ defined in \pref{def:glc} is a graph where $\V \subset \Proc \cup \Obj \cup \Sol$ and $\E \subset \V \times \V$.
A node in $\PHproc$ is a required process,
a node in $\Obj$ is an objective to reach a given process
and a node in $\Sol$ is a sets of processes required for the solving.
An objective $P \in \Obj$ is solvable if the abstractions of bounce sequences $\aBS(P) \in \Sol$ (\pref{def:aBS}) can be reached (\pref{eq:ESol1}), leading to several required processes (\pref{eq:ESol2}).
If $a$ is a component, the reachability of a process $a_i$ is approximated by the ability to solve all objectives $\PHobjp{a}{j}{i} \in \Obj$ for all $a_j$ in the initial context (\pref{eq:EReq});
if $a$ is a cooperative sort, the reachability of $a_i$ is directly solved by the set of the processes $\csState(a_i)$ (\pref{def:csState}) it represents (\pref{eq:EPrio}).
The resolution of an objective $P$ may a process of $\PHsort(P)$, \ie $\gCont(\PHsort(P), P) \neq \emptyset$ (\pref{def:maxCont}); in this cas, the objective is re-targeted (\pref{eq:ECont}).
\pref{eq:Vw}, \eqref{eq:Vproc} and \eqref{eq:VE} ensure that all required nodes are in $\cwV$.
Finally, as the active process of every sort may evolve, $\cwB$ is obtained iteratively by saturation with every process it contains,
\ie by overriding its initial context $\ctx$ by $\allprocs(\V, \E)$, defined by:
$$\allprocs(\V, \E) = (V \cap \Proc) \cup \{ \PHtarget(P), \PHbounce(P) \mid P \in \V \cap \Obj \}$$

@@

\begin{definition}
\label{def:glc}
  The graph of local causality $\cwB = (\Bv, \Be)$ is defined as: $\cwB = \lfp{\aB^\w_\ctx}{\myB}{\aB^\w_{\ctx \Cap \allprocs(\myB)}}$,
  where $\myB = (\cwV, \cwE)$ is the smallest graph where:
  $\cwV \subset \Proc \cup \Obj \cup \Sol$ and $\cwE \subset \cwV \times \cwV$
  so that:
  \begin{align}
    \w &\subset \cwV \label{eq:Vw} \\
    P \in \VProc &\Rightarrow \PHbounce(P) \in \cwV \label{eq:Vproc} \\
    (x, y) \in \E &\Rightarrow y \in \cwV \label{eq:VE} \\
    P \in \VObj \wedge ps \in \BS(P) &\Rightarrow (P, ps) \in \cwE \label{eq:ESol1} \\
    ps \in \VSol \wedge a_i \in ps &\Rightarrow (ps, a_i) \in \cwE \label{eq:ESol2} \\
    a \in \components \wedge a_i \in \VProc \wedge a_j \in \ctx &\Rightarrow (a_i, \PHobjp{a}{j}{i}) \in \cwE \label{eq:EReq} \\
    a \in \cs \wedge a_i \in \VProc \wedge ps \in \csState(a_i) &\Rightarrow (a_i, ps) \in \cwE \label{eq:EPrio} \\
    P \in \VObj \wedge q \in \gCont_\ctx(\PHsort(P), P) &\Rightarrow (P, \PHobj{q}{\PHbounce(P)}) \in \cwE \label{eq:ECont}
  \end{align}
\end{definition}

@@

In the graph of local causality, an edge linking a process of a cooperative sort to a solution set $ps$ is said coherent (\pref{def:coherent}) if none of the processes in $ps$ conflict with the children processes of $ps$.
Then, \pref{th:approxinf} gives a sufficient condition for the concretization of a sequence of objectives in a given context,
which is derived immediately from the graph of local causality.
A proof of this theorem is given in Annex~\ref{suppl:demoapproxinf}.
\begin{definition}[Coherent edge]
\label{def:coherent}
  An edge $(x, y) \in \cwE$ is said coherent if and only if:
  $(x, y) \in \Be \cap (\Proc \times \Sol) \Rightarrow y$ has no children process $a_j \in \Bv \cap \Proc$ so that $\exists a_i \in y$, $a_i \neq a_k$.
\end{definition}

>>>>

\section{Static Analysis}\label{sec:sa}

@@

We denote $\concr(\w)$ the set of scenarios concretising an objective sequence $\w$ in the context $\ctx$.
In \pref{def:uconcr}, we define $\uconcr(\w)$ as equal to $\concr(\w)$ if and only if $\concr(\w)$ contains scenarios starting from all states $s \subseteq \ctx$.
\pref{th:uconcr-ctx} is used to over-approximate the initial context $\ctx$.
\begin{definition}[$\uconcr: \OS \rightarrow \powerset(\Sce)$]
\label{def:uconcr}
  \[
  \uconcr(\w) = 
  \begin{cases}
    \concr(\w) & \text{if } \forall s \in \PHl, s \subseteq \ctx, \exists \delta \in \concr(\w), \delta \in \Sce(s) \\
    \emptyset & \text{else.}
  \end{cases}
  \]
\end{definition}
% 
\begin{theorem}
\label{th:uconcr-ctx}
  $\ctx \subseteq \ctx' \wedge \muconcr_{\ctx'}(\w) \neq \emptyset \Longrightarrow \muconcr_{\ctx}(\w) \neq \emptyset$.
\end{theorem}

@@

The graph of local causality $\cwB = (\V, \E)$ defined in \pref{def:glc} is a graph where $\V \subset \Proc \cup \Obj \cup \Sol$ and $\E \subset \V \times \V$.
A node in $\PHproc$ is a required process,
a node in $\Obj$ is an objective to reach a given process
and a node in $\Sol$ is a set of processes required for the solving.
An objective $P \in \Obj$ is solvable if the abstractions of bounce sequences $\aBS(P) \in \Sol$ (\pref{def:aBS}) can be reached (\pref{eq:ESol1}), thus leading to several required processes (\pref{eq:ESol2}).
If $a \in \components$, the reachability of one of its process $a_i$ is approximated by the ability to solve all objectives $\PHobjp{a}{j}{i} \in \Obj$ for all $a_j$ in the initial context (\pref{eq:EReq});
if $a \in \cs$, the reachability of $a_i$ is simply solved by the set of processes $\csState(a_i)$ (\pref{def:csState}) that it represents (\pref{eq:EPrio}).
The solving of an objective $P$ may require a process of $\PHsort(P)$, \ie $\gCont(\PHsort(P), P) \neq \emptyset$ (\pref{def:maxCont}); in this case, $P$ is re-targeted (\pref{eq:ECont}).
\pref{eq:Vw}, \eqref{eq:Vproc} and~\eqref{eq:VE} ensure that all required nodes are in $\cwV$.
Finally, as the active process of every sort may evolve, $\cwB$ is obtained by iteratively saturating with every process it contains,
\ie by overriding its initial context $\ctx$ by $\allprocs(\V, \E)$, defined by:
$$\allprocs(\V, \E) = (V \cap \Proc) \cup \{ \PHtarget(P), \PHbounce(P) \mid P \in \V \cap \Obj \}$$

@@

\begin{definition}
\label{def:glc}
  The graph of local causality $\cwB = (\Bv, \Be)$ is defined as: $\cwB = \lfp{\aB^\w_\ctx}{\myB}{\aB^\w_{\ctx \Cap \allprocs(\myB)}}$,
  where $\myB = (\cwV, \cwE)$ is the smallest graph with
  $\cwV \subset \Proc \cup \Obj \cup \Sol$ and $\cwE \subset \cwV \times \cwV$
  so that:
  \begin{align}
    \w &\subset \cwV \label{eq:Vw} \\
    P \in \VProc &\Rightarrow \PHbounce(P) \in \cwV \label{eq:Vproc} \\
    (x, y) \in \E &\Rightarrow y \in \cwV \label{eq:VE} \\
    P \in \VObj \wedge ps \in \BS(P) &\Rightarrow (P, ps) \in \cwE \label{eq:ESol1} \\
    ps \in \VSol \wedge a_i \in ps &\Rightarrow (ps, a_i) \in \cwE \label{eq:ESol2} \\
    a \in \components \wedge a_i \in \VProc \wedge a_j \in \ctx &\Rightarrow (a_i, \PHobjp{a}{j}{i}) \in \cwE \label{eq:EReq} \\
    a \in \cs \wedge a_i \in \VProc \wedge ps \in \csState(a_i) &\Rightarrow (a_i, ps) \in \cwE \label{eq:EPrio} \\
    P \in \VObj \wedge q \in \gCont_\ctx(\PHsort(P), P) &\Rightarrow (P, \PHobj{q}{\PHbounce(P)}) \in \cwE \label{eq:ECont}
  \end{align}
\end{definition}

@@

In the graph of local causality, an edge $(p, ps) \in \Proc \times \Sol$ %linking a process of a cooperative sort to a solution set $ps$
is said coherent (\pref{def:coherent}) if none of the processes in $ps$ conflict with the children processes of $ps$.
Then, \pref{th:approxinf} gives a sufficient condition for the concretization of a sequence of objectives in a given context,
which is derived immediately from the graph of local causality.
A proof of this theorem is given in Annex~\ref{suppl:demoapproxinf}.
\begin{definition}[Coherent edge]
\label{def:coherent}
  An edge $(x, y) \in \cwE$ is said coherent if and only if:
  $(x, y) \in \Be \cap (\Proc \times \Sol) \Rightarrow y$ has no children process $a_j \in \Bv \cap \Proc$ so that $\exists a_i \in y$, $a_i \neq a_k$.
\end{definition}
